---
title: "Join Lab"
author: "Dr. Shoemaker"
date: "DATA 2401"
output: html_document
---


# Joins and `dplyr` practice

Load the `"nycflights13"` package.  You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(tidyverse)
#install.packages("nycflights13")
library(nycflights13)
my_flights <- flights # store my own version
my_flights <- na.omit(my_flights) ## this dataset has data on cancelled flights too, 
                                  ## so we 'll remove those for now

```


Create a dataframe of the average arrival delays for each _destination_

```{r}
library(tidyverse)
library(nycflights13)


my_flights <- flights %>% na.omit()


avg_arr_delay <- my_flights %>%
  group_by(dest) %>%
  summarize(Avg_Arrival_Delay = mean(arr_delay, na.rm = TRUE))
print(avg_arr_delay)
```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}


joined_data <- avg_arr_delay %>%
  left_join(airports, by = c("dest" = "faa"))


print(joined_data)
View(joined_data) 

```


Note here: the column names don't match anymore! 

Which airport had the largest average arrival delay?

```{r}


joined_data <- avg_arr_delay %>%
  left_join(airports, by = c("dest" = "faa"))


worst_airport <- joined_data %>%
  arrange(desc(Avg_Arrival_Delay)) %>%
  slice(1)  


print(worst_airport)
View(worst_airport)


```



What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each _airline_, then use `left_join()` to join on the "airlines" dataframe


```{r}


avg_dep_delay <- my_flights %>%
  group_by(carrier) %>%
  summarize(Avg_Departure_Delay = mean(dep_delay, na.rm = TRUE))


joined_airlines <- avg_dep_delay %>%
  left_join(airlines, by = "carrier")


print(joined_airlines)
View(joined_airlines) 


```

Which airline had the smallest average departure delay? 



```{r}


best_airline <- joined_airlines %>%
  arrange(Avg_Departure_Delay) %>%  
  slice(1)  


print(best_airline)
View(best_airline)



```





Examples from R4DS:    

```{r}

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join: 

```{r}
flights %>% 
  semi_join(top_dest)
```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you?
```{r}
unmatched_flights <- flights %>% 
  anti_join(airports, by = c("dest" = "faa"))

print(unmatched_flights)
View(unmatched_flights)

# it tells us which dataframes have the flights with destinations that do no appear in airports.
```

What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?
```{r}
unmatched_airports <- airports %>% 
  anti_join(flights, by = c("faa" = "dest"))

print(unmatched_airports)
View(unmatched_airports)

# this tells us the dataframes that have airports that are lited in airports but are not used as destinations in flights.
```

Perform the two joins below and consider what the results are. 

```{r}
unmatched_flights <- flights %>% 
  anti_join(airports, by = c("dest" = "faa"))

print(unmatched_flights)
View(unmatched_flights)

#fligths that have landed at destinations that are not in the airports, in other words missing from airports.

unmatched_airports <- airports %>% 
  anti_join(flights, by = c("faa" = "dest"))
print(unmatched_airports)
View(unmatched_airports)

#this tells us the airports that exist in the airports but are not/never a destination in the flights.

```

